---
title: "Reproduced analysis of FARS data"
#author: "Scott Kelleher"
#date: "November 8, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr ::opts_chunk$set(echo = FALSE)
```

**Results** 


```{r message = FALSE, warning= FALSE} 
#setwd("C:/R/homework 4/Scotts project R")

library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(knitr)
load("../data/clean_fars.RData")
source("../R/fars_functions/perc_cis.R")
source("../R/fars_functions/test_trend_ca.R")
source("../R/fars_functions/test_trend_log_reg.R")

#knitr::opts_chunk$set(echo = FALSE)
```

**Percentages of drivers testing positive by drug type, sex, and year group**


```{r warning = FALSE, message = FALSE}
#load(data/clean_fars.Rmd)
clean_fars %>%
mutate(year_cat = cut(year, breaks = c(1999, 2002, 2006, 2010),
labels = c("1999-2002", "2003-2006",
"2007-2010"),
include.lowest = TRUE, right = TRUE)) %>%
filter(!is.na(sex)) %>%
group_by(drug_type, sex, year_cat) %>%
summarize(n_non_missing = sum(!is.na(positive_for_drug)),
positive_test = sum(positive_for_drug, na.rm = TRUE),
perc_positive = round(100 * positive_test / n_non_missing, 1)) %>%
select(drug_type, sex, year_cat, perc_positive) %>%
unite(sex_year_cat, sex, year_cat) %>%
spread(sex_year_cat, perc_positive) %>%
knitr::kable(col.names = c("Drug type", "F 1999-2002",
"F 2003-2006", "F 2007-2010",
"M 1999-2002", "M 2003-2006",
"M 2007-2010"))


```


**Figure 1**: Prevalence of nonalcohol drugs in fatally injured drivers by year and age group


```{r message = FALSE, warning= FALSE, echo =FALSE}

clean_fars2 <- clean_fars %>%
  filter(!is.na(positive_for_drug),
         (!is.na(agecat))) %>%
  filter(drug_type != "Alcohol") %>%
  group_by(unique_id)%>%
  summarize(positive_for_drug = any(positive_for_drug),
            agecat = first(agecat),
            year = first(year))%>%
  ungroup() %>%
  group_by(agecat, year) %>%
  summarize(perc_pos_drug = mean(positive_for_drug)*100)

ggplot(clean_fars2, aes(x= year, y = perc_pos_drug, shape = agecat))+
  geom_point(alpha = 1)+
  geom_line()+
  scale_shape_discrete(name = "Age")+
  ylab("Positive for Nonalcohol Drugs %")+
  xlab("Year")+
  scale_y_continuous(breaks = c(0 , 10 ,20 , 30))+
  theme_few()
  
```
*Figure 2: Prevalence of nonalcohol drugs in fatally injured drivers by year and drug type*
```{r message = FALSE, warning= FALSE}

clean_fars4 <- group_by(clean_fars, drug_type, year) %>%
  filter(!is.na(positive_for_drug),
         (!is.na(drug_type))) %>%
  filter(drug_type != "Alcohol") %>%
  summarize(perc_pos_drug = mean(positive_for_drug)*100) 

     
ggplot(clean_fars4, aes(x= year, y = perc_pos_drug, shape = drug_type))+
  geom_point(alpha = 1)+
  geom_line()+
  scale_shape_discrete(name = "Drug Type")+
  ylab("Positive for Drugs %")+
  xlab("Year")+
  scale_y_continuous(breaks = c(0 , 5 ,10, 15, 20 ))+
  theme_few()

```
**Figure 3: Prevalence of cannabinoid drugs in fatally injured drivers by year and age group**

```{r message = FALSE, warning= FALSE}
clean_fars5 <- group_by(clean_fars, agecat, year) %>%
  filter(!is.na(positive_for_drug),
         (!is.na(agecat))) %>%
  filter(drug_type == "Cannabinoid") %>%
  summarize(perc_pos_drug = mean(positive_for_drug)*100) 

     
ggplot(clean_fars5, aes(x= year, y = perc_pos_drug, shape = agecat))+
  geom_point(alpha = 1)+
  geom_line()+
  scale_shape_discrete(name = "Age")+
  ylab("Positive for Canninbinaoid %")+
  xlab("Year")+
  scale_y_continuous(breaks = c(0 , 5 ,10, 15, 20))+
  theme_few()


```
**Figure 3. Prevalence of drugs in fatally injured drivers for 1999 and 2010 by drug type**

```{r message = FALSE, warning= FALSE}
 clean_fars %>%
filter(year == 2010 | year == 1999) %>%
  group_by(drug_type, year) %>%
dplyr::summarize(n = sum(!is.na(positive_for_drug)),
        x = sum(positive_for_drug, na.rm = TRUE)) %>%
  dplyr::mutate(value = perc_cis(x,n)) %>%
  select(drug_type, year, value) %>%
  spread(key = year, value = value) %>% 
  rename(`Drug type` = drug_type) %>%
  knitr::kable()

```



**Statistics for testing for trend in prevalence of drugs over study years by drug type using
Cochran-Armitage trend test**


```{r message = FALSE, warning= FALSE}

drug_list <- c("Alcohol", "Nonalcohol", "Narcotic", "Depressant",
"Stimulant", "Cannabinoid", "Other")
drug_trend_tests_ca <- lapply(drug_list, test_trend_ca)
drug_trend_tests_ca <- dplyr::bind_rows(drug_trend_tests_ca) %>%
dplyr::mutate(drug = drug_list) %>%
dplyr::select(drug, Z, p.value)
drug_trend_tests_ca %>% knitr::kable()

```


**Statistics for testing for trend in prevalence of drugs over study years by drug type using Wald
test of logistic regression coefficient for "year"**
```{r message = FALSE, warning= FALSE}
drug_list <- c("Alcohol", "Nonalcohol", "Narcotic", "Depressant",
"Stimulant", "Cannabinoid", "Other")
drug_trend_tests_log_reg <- lapply(drug_list, test_trend_log_reg)
drug_trend_tests_log_reg <- dplyr::bind_rows(drug_trend_tests_log_reg) %>%
dplyr::mutate(drug = drug_list) %>%
dplyr::select(drug, Z, p.value)
drug_trend_tests_log_reg %>% knitr::kable()

```

